Networked Character Devices

### 2011-12-11 report/presentation.tex
report: put listings in their own files
I should have done this from the start but I was lazy. Now I want to
reuse them in the paper. This is another form of laziness =)

\lstloadlanguages{C}
\lstset{
  language=C,
  basicstyle=\footnotesize,
  numbers=none,
  numberstyle=\footnotesize,
  stepnumber=1,
  numbersep=5pt,
  showspaces=false,
  showstringspaces=false,
  showtabs=false,
  tabsize=4,
  captionpos=b,
  breaklines=true,
  breakatwhitespace=false,
  title=\lstname,
  frame=single,
  frameround=tttt
}

@@

\begin{frame}[fragile]{Protocol}
  \begin{columns}
    \begin{column}{0.5\textwidth}
      \begin{lstlisting}[gobble=8]
        enum FOPS {
          FOP_OPEN, FOP_RELEASE,
          FOP_READ, FOP_WRITE,
          _FOP_ERROR
        };

        struct fop_request {

          enum FOPS call;

          union {
            struct {
              int     flags;
              mode_t  mode;
            };

            size_t  count;
          };
        };
      \end{lstlisting}
    \end{column}
    \begin{column}{0.5\textwidth}
      \begin{lstlisting}[gobble=8]
        struct fop_reply {

          enum FOPS call;

          union {
            int      open;
            int      close;
            ssize_t  read;
            ssize_t  write;
          };
        };
      \end{lstlisting}
    \end{column}
  \end{columns}
\end{frame}

@@

\begin{frame}[fragile]{Hello, World!}
  \begin{columns}
    \begin{column}{0.6\textwidth}
      \begin{lstlisting}[gobble=8]
        #include <linux/kernel.h>
        #include <linux/module.h>

        static int __init nc_init(void)
        {
          /* do stuff */
          return 0;
        }

        static int __exit nc_exit(void)
        {
          /* undo stuff */
        }

        MODULE_LICENSE("GPL");

        module_init(nc_init);
        module_exit(nc_exit);

      \end{lstlisting}
    \end{column}
    \begin{column}{0.4\textwidth}
      \begin{itemize}
        \item initial setup
        \item cleanup important!
      \end{itemize}
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}[fragile]{Character Device Driver}
  \begin{lstlisting}[gobble=4]
    #include <linux/cdev.h>
    #include <linux/device.h>
    #include <linux/fs.h>

    static dev_t          nc_dev_t;
    static struct class*  nc_class;
    static struct cdev*   nc_cdev;
    static struct device* nc_device;

  \end{lstlisting}

@@

\begin{frame}[fragile]{Filesystem Handler}
  \begin{lstlisting}[gobble=4]
    #include <linux/fs.h>

    static int netchar_open(struct inode* inp, struct file* fp);
    static ssize_t netchar_read(struct file* fp, char* buffer,
                                size_t length, loff_t* offset);

    static struct file_operations nc_fops = {
      /* ... */
      .open  = nc_open,
      .read  = nc_read,
    }
  \end{lstlisting}

@@

\begin{frame}[fragile]{Server (\texttt{read()})}
  \begin{lstlisting}[gobble=4]
    struct fop_request req;
    struct fop_reply   rep;
    while (1) {
      err = read(sd, &req, sizeof(req));
      switch(req.call) {
      case FOP_READ:
        payload = realloc(payload, req.count);
        rep.read = read(fd, payload, req.count);
        if (rep.read < 0)
          rep.read = -errno;
        break;
      }
      err = write(sd, &rep, sizeof(rep));
      if (rep.call == FOP_READ && rep.read > 0)
        err = write(sd, payload, rep.read)
    }
  \end{lstlisting}
\end{frame}

@@

\begin{frame}[t,fragile]{\texttt{udev}}
  \begin{lstlisting}[gobble=4,title=no way to set permissions within kernel,language=]
    ACTION=="add", SUBSYSTEM=="netchar", MODE="0666"
  \end{lstlisting}
  \begin{lstlisting}[gobble=4,title=cannot retroactively apply input attributes,language=]
    $ udevadm info -qall -n /dev/input/event8
    P: /devices/virtual/input/input8/event8
    N: input/event8
    E: DEVNAME=/dev/input/event8
    E: DEVPATH=/devices/virtual/input/input8/event8
    E: ID_INPUT=1
    E: ID_INPUT_KEY=1
    E: SUBSYSTEM=input
  \end{lstlisting}
%$
\end{frame}

\begin{frame}[fragile]{\texttt{ncadmin}}
  \begin{lstlisting}[gobble=4,language=]
    usage: ncadmin command [args]

      The following commands are supported:

      add server port [name]
        server - an ip address
        port   - a port
        name   - optional name relative to /dev. if no name is
                 specified /dev/netchar/import# will be used

      rm name
        name   - the device relative to /dev to remove
  \end{lstlisting}
  \begin{itemize}
    \item<2-> does nothing!
    \item<3-> could use discovery service
    \item<4-> requires individual \texttt{cdev}s
    \item<5-> name is optional
  \end{itemize}
\end{frame}

>>>>

\lstloadlanguages{C}
\lstset{
  language=C,
  basicstyle=\footnotesize,
  numbers=none,
  numberstyle=\footnotesize,
  stepnumber=1,
  numbersep=5pt,
  showspaces=false,
  showstringspaces=false,
  showtabs=false,
  tabsize=4,
  captionpos=b,
  breaklines=true,
  breakatwhitespace=false,
  frame=single,
  frameround=tttt
}

@@

\begin{frame}{Protocol}
  \begin{columns}
    \begin{column}{0.5\textwidth}
      \lstinputlisting[lastline=19]{protocol.c}
    \end{column}
    \begin{column}{0.5\textwidth}
      \lstinputlisting[firstline=21]{protocol.c}
    \end{column}
  \end{columns}
\end{frame}

@@

\begin{frame}{Hello, World!}
  \begin{columns}
    \begin{column}{0.6\textwidth}
      \lstinputlisting[]{kmod-hello.c}
    \end{column}
    \begin{column}{0.4\textwidth}
      \begin{itemize}
        \item initial setup
        \item cleanup important!
      \end{itemize}
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}{Character Device Driver}
  \lstinputlisting[]{cdev-driver.c}

@@

\begin{frame}{Filesystem Handler}
  \lstinputlisting[]{fop-driver.c}

@@

\begin{frame}{Server (\texttt{read()})}
  \lstinputlisting[]{server.c}
\end{frame}

@@

\begin{frame}[t]{\texttt{udev}}
  \lstinputlisting[
    lastline=1,
    language={},
    title=no way to set permissions within kernel,
  ]{udev.txt}
  \lstinputlisting[
    firstline=2,
    language={},
    title=cannot retroactively apply input attributes
  ]{udev.txt}
%$
\end{frame}

\begin{frame}{\texttt{ncadmin}}
  \lstinputlisting[]{ncadmin.txt}
  \begin{itemize}
    \item<2-> does nothing!
    \item<3-> could use discovery service
    \item<4-> requires individual \texttt{cdev}s
    \item<5-> name is optional
  \end{itemize}
\end{frame}
